\myTitle

### 2024-05-31 ckmr-sim-evaluation.Rmd
Add comparison of CKMR logls with all SNPs vs...
just the top SNPs from each amplicon.

```{r}
if(exists("snakemake")) {
  input_list <- snakemake@input
  output_list <- snakemake@output
} else {
  input_list <- list(
    final_baseline = "data/subRoSA_baseline_rubias_FEB2024_w_mixture_revised.csv",
    lfar_genos = "data/metagenos_hap2col_lfar_all_samps.csv",
    pop_labels = "inputs/reference-collection-names.csv",
    locus_info = "inputs/ChinookFullPanelLociTable_Feb2024Updated.xlsx"
  )
  output_list <- list(
    ckmr_figure = "results/ckmr-sim-eval/fpr-fnr-figure.pdf",
    ckmr_figure_crop = "results/ckmr-sim-eval/fpr-fnr-figure-crop.pdf",
    ckmr_figure_tex = "tex/images/fpr-fnr-figure-crop.pdf"
  )
}

@@

>>>>

```{r}
if(exists("snakemake")) {
  input_list <- snakemake@input
  output_list <- snakemake@output
} else {
  input_list <- list(
    final_baseline = "data/subRoSA_baseline_rubias_FEB2024_w_mixture_revised.csv",
    lfar_genos = "data/metagenos_hap2col_lfar_all_samps.csv",
    pop_labels = "inputs/reference-collection-names.csv",
    locus_info = "inputs/ChinookFullPanelLociTable_May2024Updated.xlsx"
  )
  output_list <- list(
    ckmr_figure = "results/ckmr-sim-eval/fpr-fnr-figure.pdf",
    ckmr_figure_crop = "results/ckmr-sim-eval/fpr-fnr-figure-crop.pdf",
    ckmr_figure_tex = "tex/images/fpr-fnr-figure-crop.pdf",
    ckmr_comp = "results/ckmr-sim-eval/cmkr-comp-figure.pdf",
    ckmr_comp_crop = "results/ckmr-sim-eval/cmkr-comp-figure-crop.pdf",
    ckmr_comp_tex = "tex/images/cmkr-comp-figure-crop.pdf"

  )
}

@@

# Logl Ratio Plot Comparisons

Neil suggested, and Carlos and I agreed, that one way to see they improvement in
relationship inference obtained by using multiallelic microhaps is to look at the
separation of the different relationship log-likelihood ratios when using just the
best (MAF closest to 0.5) SNP from each locus, or when using all the SNPs/alleles.
We deciced to do this for the hatchery populations in California, as well as the
winter run.  I am going to implement this in a few functions, each of which operate
on the one of the ckmr_sim_nests.

```{r}
#' function to get just the best SNP out of each microhap
#' 
#' This works on a single collection at a time.
#' @param X a tibble like that nested in a single row of ckmr_sim_nests
#' @return Returns a tibble like X, but with just the best SNP from
#' each amplicon (note that it might also drop some monomorphic amplicons).
retain_top_snp <- function(X) {
  # first, get the microhaps in long format
  tmp <- X %>%
    select(-(repunit:sample_type)) %>%
    pivot_longer(
      -indiv,
      names_to = c("Locus", "gene_copy"),
      values_to = "Allele",
      names_pattern = "^(.*)_([12])$"
    ) %>%
    filter(!is.na(Allele))

  # then explode it out to individual SNPs
  exploded <- tmp %>%
    mutate(
      snp_num = map(Allele, function(x) 1:nchar(x)),
      snp_alle = map(Allele, function(x) strsplit(x, "")[[1]])
    ) %>%
    unnest(cols = c(snp_num, snp_alle))

  # now, get the highest maf snp at each locus in a tibble
  highest_mafs <- exploded %>%
    count(Locus, snp_num, snp_alle) %>%
    group_by(Locus, snp_num) %>%
    mutate(af = n / sum(n)) %>%
    filter(af <= 0.5) %>%
    ungroup() %>%
    arrange(Locus, desc(af)) %>%
    group_by(Locus) %>%
    slice(1) %>%
    ungroup()

  # then pull those out of the "exploded" variable with a semi_join
  # and put them back into wide format, left join it onto the meta
  # data
  ret <- exploded %>%
    semi_join(highest_mafs, by = join_by(Locus, snp_num)) %>%
    select(indiv, Locus, gene_copy, snp_alle) %>%
    pivot_wider(
      names_from = c(Locus, gene_copy),
      values_from = snp_alle,
      names_sep = "_"
    ) %>%
    left_join(
      X %>% select(indiv, repunit, collection, sample_type),
      .,
      by = join_by(indiv)
    )

  ret

So, that is how we pick out the top SNP.  But now we also need a function that is
going to give us the set of all the simulated logl-ratios that we need (whether it
is all the microhap alleles or just the top SNPs).  We can build off of what
we did in a function above.
```{r}
get_logl_rats <- function(X, GC2 = gen_coords) {

  # Simulate some Q_ij's. Suppress output messages that the CKMRsim package
  # produces
  Q_link <- ckmr_simQ(C) #%>% suppressMessages()

  # now extract the logl_ratios for PO vs U and return those
  extract_logls(Q_link, numer = c(PO = 1), denom = c(U = 1))
}
```

And, with those two functions, it would appear that we have eveything we need to get the
simulation results we will want.  Once again we will mclapply over the populations.

```{r}
# collections / groups we want to do
pops <- c("ck-FRH", "ck-IGH", "ck-SRW", "ck-TRH")

# let's put the microhaps and the best SNPs together in a long format.
input_tibbles <- bind_rows(
  ckmr_sim_nests %>%
    filter(ckmr_sim_group %in% pops) %>%
    mutate(condition = "microhaps", .before = ckmr_sim_group),
  ckmr_sim_nests %>%
    filter(ckmr_sim_group %in% pops) %>%
    mutate(tops = map(data, retain_top_snp)) %>%
    mutate(data = tops) %>%
    select(-tops) %>%
    mutate(condition = "top_snps", .before = ckmr_sim_group)
)

Qsims <- mclapply(input_tibbles$data, get_logl_rats, mc.cores = 8)

AllSims <- tibble(
  condition = input_tibbles$condition,
  ckmr_sim_group = input_tibbles$ckmr_sim_group,
  Qsims = Qsims
) %>% 
  unnest(Qsims) %>%
  select(condition, ckmr_sim_group, true_relat, logl_ratio)
```

Now, we can plot all of those.  Let's see if we can get all four onto one page easily enough.
```{r}
asg <- ggplot(AllSims, aes(x = logl_ratio, fill = true_relat)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~ ckmr_sim_group + condition, ncol = 1) +
  theme_bw() +
  labs(fill = "True\nRelationship") +
  xlab("PO vs U Log-likelihood Ratio")

# this fails for some reason (ggsave never returns)
# ggsave(asg, path = "boing.png", height = 15, width = 9)

# So, instead I just print the thing in the Rstudio window and export it
# at size 13" high by 10" wide. Then I moved it to where it belongs in the tex directory.
asg
```

### 2024-05-31 tex/supplemental-figures.tex
Add comparison of CKMR logls with all SNPs vs...
just the top SNPs from each amplicon.

>>>>

%%%%%%%%%%%%%%%%%%%%%% 
\begin{figure}
\begin{center}
\includegraphics[width=0.9\textwidth]{images/cmkr-comp-figure-crop.pdf}
\end{center}
\caption[Comparison of microhaplotypes vs.~best SNP in each amplicon]{\footnotesize 
A comparison of the distribution of log-likelihood ratios when using all the alleles at each
amplicon typed as microhaplotypes versus using just a single (the most informative) SNP from each
amplicon.  Results shown for four hatchery collections in California.  The density plots show
the distribution of log-likelihood ratios for Parent-Offspring vs Unrelated for four different relationships:
PO = parent offspring; FS = full-sibling; HS = Half sibling; AN = avuncular (i.e., aunt-niece, etc.)
The distributions for AN and HS overlap completely.  Note that the overlap between FS and PO occurs
for the PO vs U likelihood ratio, but is nearly eliminated with the PO vs FS likelihood ratio, allowing these
two relationships to be resolved accurately.}
\label{fig:ckmr-comp}
\end{figure}
%%%%%%%%%%%%%%%%%%%%
